---
title: "R Markdown Report"
author: "Max Tölle"
output:
  html_document: default
  pdf_document: default
---

# Script and data overview for Jinggang et al. 2024

In this Script, the processing of NPP data and of FOROM output data is compiled

1.  NPP data 1.1 NPP data import 1.2 data preparation 1.3 five-year rolling average and percentage calculation 1.4 NPP figures

2.  FOROM data 2.1 import Price data 2.2 import Quantity data 2.3 merge data by continent 2.4 FOROMO figures

3.  Random Forest Analysis

**!!Please adjust directories to your local directories!!**

## 1. NPP data

NPP model outputs were collected by Jinfeng Chang (College of Environmental and Resource Sciences, Zhejiang University, Hangzhou, China) and are currently stored at <https://www.pik-potsdam.de/~reyer/ISIMIP2b_forest_csv_2022-11-25/>.

The data were processed as followed:

Necessary libraries

```{r message=FALSE, warning=FALSE}

rm(list = ls())
options(scipen = 999)

library(ncdf4)
library(terra)
library(itsadug)
library("RColorBrewer")
library(reshape2)
library(tidyverse); library(fs); library(readxl)
library("scales")
library(ggplot2)
library(plyr)
library(dplyr)
library(purrr)
library(forestmangr)
library(ggpubr)
library(tidyterra)
library(zoo)
```

Files and Script preparation:

```{r message=FALSE, warning=FALSE}
FOROM_regions <- read.table("C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/FOROM_regions.txt", fill = TRUE) #names of the 50 regions
```

### 1.1 NPP data import

import NPP data for the GVMs CARAIB, LPJ-GUESS, LPJmL, ORCHIDEE

```{r message=FALSE, warning=FALSE}
#create listing (fs_path object) of relevant files to read
excel_files <- dir_ls(path =  "C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/NPP_new", regexp = "*.csv") 

#read all files and store their name in `source` variable
excel_files <- excel_files %>% 
  set_names() %>% 
  map_dfr(read_csv, .id = "source")

excel_files$source <- as.character(excel_files$source) 

#separate source into columns to get vegmodel, GCM and RCP
excel_files <- excel_files %>% 
  mutate(source = str_remove(source, "(.*)ISIMIP2b_")) %>% 
  separate(source, into = c("variable","forest", "vegmodel", "GCM", "rcp" ), sep = "_") %>% 
  subset(select = -c(forest,variable))
  
#remove extension from rcp column 
excel_files <- excel_files %>% 
  mutate(rcp = str_remove(rcp, ".csv")) 

names(excel_files)[4] <- "year"

#calculate actual year 
excel_files <- excel_files %>% 
  mutate(year = ifelse(rcp == "historical", year + 1860, year)) %>% 
  mutate(year = ifelse(rcp %in% c("rcp26", "rcp60", "rcp85"), year + 2005, year))

#give FOROM region names to columns 
names(excel_files)[5:54] <- c(FOROM_regions$V2)

excel_files[excel_files == 0] <- NA
```

import NPP data for the GVM MC2 --\> right now still in extra file

```{r message = FALSE, warning=FALSE}
#create listing (fs_path object) of relevant files to read
npp_files <- dir_ls(path = "C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/MC2", regexp = "*.csv") 

#read all files and store their name in `source` variable
npp_files <- npp_files %>% 
  map_dfr(read_csv, .id = "source", skip = 1)

npp_files$source <- as.character(npp_files$source)

#separate source into columns to get vegmodel, GCM and RCP
npp_files<- npp_files %>% 
  mutate(source = str_remove(source, "(.*)MC2/")) %>% 
  separate(source, into = c( "GCM", "rcp", "variable"), sep = "_") %>% 
  subset(select = -c(variable))
  
npp_files <- npp_files[-54]

#give FOROM region names to columns 
names(npp_files)[4:53] <- c(FOROM_regions$V2)

npp_files$vegmodel <- c("MC2")
npp_files <- npp_files %>% 
  relocate(vegmodel)

npp_files <- npp_files %>% 
  mutate(rcp = ifelse(year <= 2005, "historical", rcp))
  
#no data for NA_ROC is available for MC2
npp_files$NA_ROC <- NA 
```

merge NPP data of vegetation models

```{r}
npp.data <- rbind(npp_files, excel_files) 
rm(excel_files,npp_files)
```

### 1.2 data preparation

Subset the data to GVMs RCPs and time that is used:

GVMs: CARAIB, LPJmL,LPJ-GUESS, ORCHIDEE & MC2

GCMs: gfdl-esm2m, hadgem2-es, ipsl-cm5a-lr, miroc5

time: 1964 - 2099

RCPs: rcp26, rcp60, rcp85

```{r}
# CLM45 is not used in this study! 
npp.data <- subset(npp.data, npp.data$vegmodel != "CLM45") 

# subset to time of concern
npp.data <- subset(npp.data, npp.data$year %in% c(1964:2099)) 

#CARAIB did not provide data for rcp85
npp.data <-npp.data %>%
  filter(!(GCM=="gfdl-esm2m" & vegmodel== "CARAIB" & rcp =="rcp85" )) %>%
  filter(!(GCM=="hadgem2-es" & vegmodel== "CARAIB"& rcp =="rcp85" )) %>%
  filter(!(GCM=="ipsl-cm5a-lr" & vegmodel== "CARAIB" & rcp =="rcp85")) %>%
  filter(!(GCM=="miroc5" & vegmodel== "CARAIB"& rcp =="rcp85" )) 
  
  
```

overview of NPP data:

```{r}
head(npp.data, 10) #just as an example
```

transform data from wide to long, so that each a new region column gets created:

```{r}
npp.data <- melt(npp.data, id.vars = c("vegmodel", "GCM", "rcp", "year"))

colnames(npp.data)[colnames(npp.data) == "variable"] <- "region"

str(npp.data) # region now as a factor-column 

```

### 1.3 five-year rolling average and percentage to baseline calculation

5-year rolling averages for NPP data

```{r}
rolling_avg_window <- 5 # take the last 5 years for the calculation of mean 

rolling.data <- npp.data %>%
  group_by(vegmodel, GCM, rcp,region) %>%
  mutate(RollingAverage = rollmean(value, k = rolling_avg_window, align = "right", fill = NA))
```

Calculate the average RollingAverage between 1980-2005 for each RCP and Region

```{r, message=FALSE}
rolling.baseline <- rolling.data %>%
  filter(year >= 1980 & year <= 2005) %>%
  group_by(vegmodel, GCM, rcp,region) %>%
  summarize(baseline = mean(RollingAverage, na.rm = TRUE)) %>% 
  subset(select = -c(rcp))
```


Calculate rolling average of years after 2005


```{r}
rolling.data <- rolling.data %>%
  filter(year > 2005) %>%
  group_by(vegmodel, GCM, rcp,region, year) %>%
  summarize(value = mean(RollingAverage, na.rm = TRUE))
```


Join the average RollingAverage of the baseline period back to the original dataframe

```{r}
rolling.data <- rolling.data %>%
  left_join(rolling.baseline, by = c("vegmodel","GCM","region"))
```

Calculate the percentage increase/decrease of the actual rolling average in NPP compared to the baseline period

```{r}
rolling.data$percentage <- (rolling.data$value - rolling.data$baseline) / rolling.data$baseline *100

#filter for the 5-year period 
dt <- rolling.data %>% 
  filter(year %in% c(2020,2025,2030,2035,2040,2045,2050,2055,2060,2065,2070))%>%
  select(vegmodel, GCM,rcp, region,year,value,baseline,percentage) 
```

This dataset "dt" is used for the simulation in the FOROM model with only minor adjustments. After the model was run, the output is again processed in R!

### 1.4 NPP figures

**Figure SOM2 - historic and future NPP trends**

```{r}
world <- vect("C:/Users/maxto/Desktop/HIWI-Arbeit/FOROM_paper/data/world_shp/World_regions_50_new.shp")
country_code <- read.csv("C:/Users/maxto/Desktop/HIWI-Arbeit/FOROM_paper/data/country_code.csv")
world <- merge(world, country_code, by.x = "RegionNam", by.y = "country", all.x = TRUE)

vegmodel <- c("CARAIB", "LPJmL", "LPJ-GUESS", "MC2", "ORCHIDEE")
rcp <- c("historic", "rcp26", "rcp60", "rcp85")
```

preparation of data

```{r, warning = FALSE}

#### separate data in future and historic 
  npp_hist <- subset(npp.data, npp.data$year %in% c(1980:2005))
  npp_fut <- subset(npp.data, npp.data$year %in% c(2041:2070))

  
#-----------------------#
#### linear models ####
#-----------------------#

#### historic 
  
lm_hist <- npp_hist %>% 
    dplyr::group_by(year, vegmodel, rcp, region) %>%
    dplyr::summarise_at(vars(-c(GCM)), funs(mean(value, na.rm=TRUE))) %>%
    na.omit() %>%
    dplyr::group_by(region, vegmodel, rcp) %>% 
    lm_table(value ~ year, df = .)
  

  lm_hist <-with(lm_hist,  lm_hist[order(as.character(region)) , ])
  lm_hist <- aggregate(lm_hist$b1, by=list( vegmodel = lm_hist$vegmodel, region = lm_hist$region), FUN=mean) 
  
  lm_hist$rcp <- "historic"
  lm_hist$b1 <- lm_hist$x
  lm_hist <- lm_hist[-3]
  
  
#### future 
  
  lm_fut <- npp_fut %>% 
    dplyr::group_by(year, vegmodel, rcp, region) %>%
    dplyr::summarise_at(vars(-c(GCM)), funs (mean(value, na.rm=TRUE)))%>%
    na.omit() %>%
    dplyr::group_by(region, vegmodel, rcp) %>% 
    lm_table(value ~ year, df = .)
  
  
  lm_fut <-with(lm_fut,  lm_fut[order(as.character(region)) , ])
  
#### combine 
  
  lm_data <- bind_rows(lm_fut, lm_hist)
  
  
  fut.world <- merge(world, lm_data, by.x = "ISO", by.y = "region", all.x = TRUE)

```

creating the figure

```{r}
 
  plot_list1 <- list()
  nested_list <- list()

  for (i in 1:length(vegmodel)) {
    for (j in 1:length(rcp)) {
      if(vegmodel[i]== "CARAIB" & rcp[j] == "rcp85"){plot_list1[[j]] <- NA} 
      else {
        
        plot  <- ggplot(data = subset(fut.world, fut.world$vegmodel == vegmodel[i] & fut.world$rcp == rcp[j]), aes(fill = b1))+ 
          geom_sf()+ 
          coord_sf(crs = 4326, expand = TRUE)+
          theme_classic()+ 
          labs(title = "")+
          ylab(expression(paste("NPP [", ~frac(kg,m^2 ~ s),"]")))+
          scale_fill_gradientn(
            name = expression(paste("NPP [", ~frac(kg,m^2 ~ s),"] per year")),
            colors=c("darkred","white","forestgreen"),
            values=rescale(c(-2.5,0,6)),
            limits=c(-2.5,6),
            oob=squish
          )+
          theme(panel.spacing.y = unit(10, "lines"), 
                axis.line.x =   element_line(colour = "black"),
                axis.line.y = element_line(color="black"),
                axis.text.x = element_text(size = 10, colour = "black"),
                #axis.ticks.x = element_blank(),
                legend.title = element_text(size = 14, face = "bold"),
                legend.text = element_text(size = 12, face = "bold"),
                legend.key.height  = unit(4, "cm"))
        
        plot_list1[[j]] <- plot
      }
    }
    nested_list[[i]] <- plot_list1
  }
  
  combined.plot <- ggarrange(plotlist = c(nested_list[[1]],nested_list[[2]],nested_list[[3]],nested_list[[4]],nested_list[[5]]), labels = list("historic", "   rcp26", "   rcp60","   rcp85"), legend = "right", ncol = 4, nrow = 5, common.legend = TRUE, widths = c(0.25,0.25,0.25,0.25), hjust = -2.5, font.label = list(size = 16, face = "bold", color ="black"))
  
  
#jpeg(file="C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/figures/npp_world_trend_new.jpeg", width = 1400, height = 800, quality = 1000)
  print(annotate_figure(combined.plot, 
                        left = text_grob("              ORCHIDEE                          MC2                            LPJmL                        LPJ-GUESS                       CARAIB", rot = 90, hjust = 0.55,
                                         color = "black", face = "bold", size = 16)))
  
#while (!is.null(dev.list()))  dev.off()
```

**Figure SOM3 - Future NPP percentage compared to baseline 1980-2005**

```{r}
rcp <- c("rcp26", "rcp60", "rcp85")
```

```{r}
npp_hist_mean <- npp_hist %>% 
  group_by(vegmodel, rcp, region) %>%
  summarise_at(vars(-c(GCM, year)), funs (mean(., na.rm=TRUE))) 


npp_hist_mean <- with(npp_hist_mean, npp_hist_mean[order(as.character(region)) , ])


npp_fut_mean <- npp_fut %>% 
  group_by(vegmodel, rcp, region) %>%
  summarise_at(vars(-c(GCM, year)), funs (mean(., na.rm=TRUE))) 

npp_fut_mean <- with(npp_fut_mean, npp_fut_mean[order(as.character(region)) , ])

npp_mean <- merge(npp_hist_mean[, c("vegmodel", "region", "value")], npp_fut_mean[, c("vegmodel", "rcp", "region", "value")], by.x = c("vegmodel", "region"), by.y =c("vegmodel", "region"), all.y = TRUE) 

colnames(npp_mean) <- c("vegmodel","region","hist", "rcp", "fut")

npp_mean$perc <- (npp_mean$fut-npp_mean$hist)/npp_mean$hist*100

# bind values to world vector 

mean.world <- merge(world, npp_mean, by.x = "ISO", by.y = "region", na.rm = FALSE)

mean.world <- mean.world[!grepl("NaN", mean.world$perc),]
```

figure creation

```{r}

plot_list1 <- list()
nested_list <- list()

for (i in 1:length(vegmodel)) {
  for (j in 1:length(rcp)) {
    plot  <- ggplot(subset(mean.world, mean.world$vegmodel == vegmodel[i] & mean.world$rcp == rcp[j]), aes(fill = perc))+ 
      geom_sf()+ 
      coord_sf(crs = 4326)+
      #facet_wrap(vegmodel~rcp)+ 
      theme_classic()+ 
      labs(title = "")+
      ylab(expression(paste("Percentage [%]")))+
      scale_fill_gradientn(
        name = "Percentage [%]",
        colors=c("darkred","white","forestgreen"),
        values=rescale(c(-25,0,100)),
        limits=c(-25,100),
        oob=squish
      )+
      theme(panel.spacing.y = unit(10, "lines"), 
            axis.line.x =   element_line(colour = "black"),
            axis.text.x = element_text(size = 12, colour = "black"),
            #axis.ticks.x = element_blank(),
            legend.title = element_text(size = 14, face = "bold"),
            legend.text = element_text(size = 12, face = "bold"),
            legend.key.height  = unit(4, "cm"))
    
    plot_list1[[j]] <- plot
  }
  nested_list[[i]] <- plot_list1
}

nested_list[[1]][[3]] <- NA # for CARAIB rcp85

combined.plot <- ggarrange(plotlist = c(nested_list[[1]],nested_list[[2]],nested_list[[3]],nested_list[[4]],nested_list[[5]]), labels = list("rcp26", "rcp60","rcp85"), legend = "right", ncol = 3, nrow = 5, common.legend = TRUE, widths = c(0.3,0.3,0.3), hjust = -4, font.label = list(size = 16, face = "bold", color ="black"))

jpeg(file="npp_world_menachange.jpeg", width = 1200, height = 800, quality = 100)
print(annotate_figure(combined.plot, 
                      left = text_grob("              ORCHIDEE                          MC2                            LPJmL                        LPJ-GUESS                       CARAIB", rot = 90, hjust = 0.55,
                                       color = "black", face = "bold", size = 16)))
dev.off()
```

**Figure SOM4 - CV agreement of models**

build cv

```{r, warning=FALSE}
npp_fut_sd <- npp_fut %>% 
  group_by(vegmodel, rcp, year, region) %>%
  summarise_at(vars(-c(GCM)), funs(sd(value, na.rm=TRUE))) 


npp_fut_mean <- npp_fut %>% 
  group_by(vegmodel, rcp, year, region) %>%
  summarise_at(vars(-c(GCM)), funs(mean(value, na.rm=TRUE))) 


npp_fut_cv <- merge(npp_fut_mean[,c("vegmodel", "rcp", "region", "year", "value")], npp_fut_sd[,c("vegmodel", "rcp", "region", "year", "value")], by.x = c("vegmodel", "rcp", "region", "year"), by.y = c("vegmodel", "rcp", "region", "year"))  


npp_fut_cv$cv <- npp_fut_cv$value.y/npp_fut_cv$value.x *100


npp_fut_cv <- npp_fut_cv %>% 
  group_by(vegmodel, rcp, region) %>%
  summarise_at(vars(-c(year)), funs(mean(cv, na.rm=TRUE))) 



cv.world <- merge(world, npp_fut_cv, by.x = "ISO", by.y = "region", na.rm = FALSE)

cv.world <- cv.world[!grepl("NaN", cv.world$value),]

```

figure creation

```{r}
plot_list1 <- list()
nested_list <- list()

for (i in 1:length(vegmodel)) {
  for (j in 1:length(rcp)) {
    plot  <- ggplot(subset(cv.world, cv.world$vegmodel == vegmodel[i] & cv.world$rcp == rcp[j]), aes(fill = cv))+ 
      geom_spatvector()+ 
      coord_sf(crs = 4326)+
      theme_classic()+ 
      labs(title = "")+
      ylab("Mean CV")+
      scale_fill_gradientn(
        name = "Mean future CV",
        colors=c("white", "darkred"),
        values=rescale(c(0,100)),
        limits=c(0,100),
        oob=squish
      )+
      theme(panel.spacing.y = unit(10, "lines"), 
            axis.line.x =   element_line(colour = "black"),
            axis.text.x = element_text(size = 12, colour = "black"),
            #axis.ticks.x = element_blank(),
            legend.title = element_text(size = 14, face = "bold"),
            legend.text = element_text(size = 12, face = "bold"),
            legend.key.height  = unit(4, "cm"))
    
    plot_list1[[j]] <- plot
  }
  nested_list[[i]] <- plot_list1
}

nested_list[[1]][[3]] <- NA

combined.plot <- ggarrange(plotlist = c(nested_list[[1]],nested_list[[2]],nested_list[[3]],nested_list[[4]],nested_list[[5]]), labels = list("rcp26", "rcp60","rcp85"), legend = "right", ncol = 3, nrow = 5, common.legend = TRUE, widths = c(0.3,0.3,0.3), hjust = -4, font.label = list(size = 16, face = "bold", color ="black"))


#jpeg(file="cv_world_npp_fut.jpeg", width = 1200, height = 800, quality = 100)
print(annotate_figure(combined.plot, 
                      left = text_grob("              ORCHIDEE                          MC2                            LPJmL                        LPJ-GUESS                       CARAIB", rot = 90, hjust = 0.55,
                                       color = "black", face = "bold", size = 16)))
#dev.off()


```

## 2. FOROM data

GAMS need to be installed on your computer in order to convert the .gdx-file in a readable R table. You can download it here: <https://www.gams.com/download/>

Loading necessary packages:

```{r message=FALSE, warning=FALSE}
library(scales)
library(readr)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(zoo)
library(tidyr)
library(sf)
library(viridis)
library(ggtext)
library(data.table)
# install.packages('gdxrrw')
# # if this is not
# working try
# install_github('GAMS-dev/gdxrrw/gdxrrw')
library(gdxrrw)
library(reshape2)
```

set igdx to local file of GAMS

```{r, message = FALSE}
igdx("C:/GAMS/46")

```

### 2.1 import Price (PS_all) and percentage change calculation

1.) Import dataset

```{r message=FALSE, warning=FALSE}
price_df <- data.table(rgdx.param("C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/merged.gdx","PS_ALL"))
colnames(price_df) <- c("SSP","Scenario", "Product","Region", "Year","Value")
price_df$Scenario <- as.character(price_df$Scenario)
price_df$Scenario <- ifelse(grepl("^[[:lower:]a-z]",price_df$Scenario),paste0("MC2_",price_df$Scenario),price_df$Scenario)
```

2.) filter the dataframe!

```{r message = FALSE, warning = FALSE}
price_df <- price_df %>%
  filter(Year %in% c("2020","2050", "2070"))  %>%
  separate(Scenario, into = c("DGVM", "GCM", "RCP"), sep = "_") #%>%

```

3.) take the mean over GCMs and calculate the percentage change

```{r}
##Summarize for mean and SD over GCMs 
percentage_changes_price <- price_df %>%
group_by(SSP, Product,Region, Year,DGVM, RCP) %>%
summarize(AverageValue = mean(Value,na.rm = TRUE), 
          SD_Value = sd(Value, na.rm = TRUE),.groups = "drop")

#transform data set 
percentage_changes_price <- percentage_changes_price %>%
  filter(Year %in% c("2020", "2070")) %>%
  pivot_wider(id_cols = c("SSP", "Product", "Region", "DGVM","RCP"),
              names_from = Year,
              values_from = c(AverageValue, SD_Value))

#calculate percentage
percentage_changes_price$percentage <- (percentage_changes_price$AverageValue_2070-percentage_changes_price$AverageValue_2020)/percentage_changes_price$AverageValue_2020*100


```

### 2.2 import Quantity (QS_all) and percentage change calculation

1.) Import dataset

```{r}
quantity_df <- data.table(rgdx.param("C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/merged.gdx","QS_ALL"))
colnames(quantity_df) <- c("SSP","Scenario", "Product","Region", "Year","Value")
quantity_df$Scenario <- as.character(quantity_df$Scenario)
quantity_df$Scenario <- ifelse(grepl("^[[:lower:]a-z]",quantity_df$Scenario),paste0("MC2_",quantity_df$Scenario),quantity_df$Scenario)
```

2.) filter the dataframe!

```{r warning=FALSE}
quantity_df <- quantity_df %>%
  filter(Year %in% c("2020", "2050","2070"))  %>%
  separate(Scenario, into = c("DGVM", "GCM", "RCP"), sep = "_") #%>%

```

3.) take the mean over the GCMs and calculate the percentage change from 2070 to 2020

```{r}
percentage_changes_quantity <- quantity_df %>%
group_by(SSP, Product,Region, Year,DGVM, RCP) %>%
summarize(AverageValue = mean(Value,na.rm = TRUE), 
          SD_Value = sd(Value, na.rm = TRUE),.groups = "drop")

percentage_changes_quantity <- percentage_changes_quantity %>%
  filter(Year %in% c("2020", "2070")) %>%
  pivot_wider(id_cols = c("SSP", "Product", "Region", "DGVM","RCP"),
              names_from = Year,
              values_from = c(AverageValue, SD_Value))

#calculate percentage
percentage_changes_quantity$percentage2070 <- (percentage_changes_quantity$AverageValue_2070-percentage_changes_quantity$AverageValue_2020)/percentage_changes_quantity$AverageValue_2020*100


```

### 2.3 Data merging by continent

```{r}
region_to_rcode <- c(
  "Australia" = "Oceania",
  "Rest of Oceania" = "Oceania",
  "Chile" = "South America",
  "China" = "Asia",
  "France" = "Europe",
  "India" = "Asia",
  "Japan" = "Asia",
  "Russian Federation" = "Asia",
  "Rest of Africa" = "Africa",
  "Rest of Europe" = "Europe",
  "Rest of North America" = "North America",
  "Mexico" = "North America",
  "Pacific Coast" = "North America",
  "North Central" = "North America",
  "South Central" = "North America",
  "Argentina" = "South America",
  "Belarus" = "Europe",
  "Estonia" = "Europe",
  "Malaysia" = "Asia",
  "Nigeria" = "Africa",
  "Portugal" = "Europe",
  "Slovakia" = "Europe",
  "Spain" = "Europe",
  "Turkey" = "Asia",
  "United Kingdom" = "Europe",
  "Viet Nam" = "Asia",
  "New Zealand" = "Oceania",
  "Brazil" = "South America",
  "Rest of South America" = "South America",
  "Finland" = "Europe",
  "Germany" = "Europe",
  "Indonesia" = "Asia",
  "Poland" = "Europe",
  "Sweden" = "Europe",
  "Rest of Asia" = "Asia",
  "Rest of Central America" = "Central America",
  "Canada" = "North America",
  # "United States" = "North America",
  "Rocky Mountain" = "North America",
  "North East" = "North America",
  "South East" = "North America",
  "Austria" = "Europe",
  "Belgium" = "Europe",
  "Latvia" = "Europe",
  "Myanmar" = "Asia",
  "Norway" = "Europe",
  "Romania" = "Europe",
  "South Africa" = "Africa",
  "Thailand" = "Asia",
  "Ukraine" = "Europe",
  "Uruguay" = "South America"
)


  percentage_changes_price$continent <- region_to_rcode[as.character(percentage_changes_price$Region)]
  percentage_changes_quantity$continent <- region_to_rcode[as.character(percentage_changes_quantity$Region)]

```

```{r}
continent_price <- percentage_changes_price %>% 
  group_by(SSP, Product,DGVM,RCP, continent) %>% 
  summarize(sum2020 = sum(AverageValue_2020, na.rm = TRUE),
            sum2070 = sum(AverageValue_2070, na.rm = TRUE))

continent_price$percentage <- (continent_price$sum2070 - continent_price$sum2020) / continent_price$sum2020 * 100
```

```{r}
continent_quantity <- percentage_changes_quantity %>% 
  group_by(SSP, Product,DGVM,RCP, continent) %>% 
  summarize(sum2020 = sum(AverageValue_2020, na.rm = TRUE),
            sum2070 = sum(AverageValue_2070, na.rm = TRUE))

continent_quantity$percentage <- (continent_quantity$sum2070 - continent_quantity$sum2020) / continent_quantity$sum2020 * 100
```

### 2.4 FOROM figures

Libraries:

```{r, warning=FALSE}
library(ggplot2)
library(data.table)
library(plyr)
library(dplyr)
library(readxl)
library(RColorBrewer)
#install.packages("ggpattern")
library(ggpattern)
library(tidyr)
library(ggnewscale)
library(ggpubr)
library(cowplot)
```

**Figure 1 - Quantity and Price**

```{r}
combined_data <- left_join(price_df, quantity_df, by = c("SSP","Product", "Region", "Year", "DGVM", "RCP", "GCM")) %>%
  rename(Price = Value.x, Quantity = Value.y) 

combined_data$continent <- region_to_rcode[as.character(combined_data$Region)]

baseline_plot <- subset(combined_data, GCM == "baseline" & Year %in% c("2020","2070") & Product %in% c("IRC", "IRNC", "SWC", "SWNC") & Region %in% c("Europe", "Asia", "Africa", "Oceania", "South America", "North America", "Central America", "WORLD"))

data_plot <- subset(combined_data, GCM != "baseline" & Year == "2070" & Product %in% c("IRC", "IRNC", "SWC", "SWNC") & Region %in% c("Europe", "Asia", "Africa", "Oceania", "South America", "North America", "Central America", "WORLD"))

data_plot <- left_join(data_plot , baseline_plot %>% select(SSP,Product, Region, Year,Price, Quantity) %>% filter(Year == 2070), by = c("SSP", "Region", "Year", "Product")) %>% rename(Price = Price.x, Price_baseline = Price.y, Quantity = Quantity.x, Quantity_baseline = Quantity.y)

data_plot$Quantity_prct <- (data_plot$Quantity - data_plot$Quantity_baseline) / data_plot$Quantity_baseline * 100
data_plot$Price_prct <- (data_plot$Price - data_plot$Price_baseline) / data_plot$Price_baseline * 100

```

Quantity

```{r, warning = FALSE, message=FALSE}

colour_2070 <- c("#C2AB44","#2353BA","#BA672B","#942323" )
colour_2020 <- c("#E3E088" , "#7DA1DB", "#E6A97E","#E06165")
Products <- as.character(unique(data_plot$Product))



actual_plot_list <- list()
baseline_plot_list <- list()
quantity_plot_list <- list()

for (i in 1:length(unique(data_plot$Region))) {
  
    act_data <-  data_plot %>% filter(Region == unique(data_plot$Region)[i], SSP == "SSP2") 
    
    actual_plot_list[[i]] <- ggplot(act_data, aes(fill = Product)) + 
      geom_bar(data = act_data %>% group_by(SSP, Region,RCP, Product, Year) %>% 
                 summarise(mean_quantity = mean(Quantity, na.rm = TRUE), 
                           SD_quantity = sd(Quantity, na.rm = TRUE),
                           mean_quantity_prct = mean(Quantity_prct, na.rm = TRUE), 
                           SD_quantity_prct = sd(Quantity_prct, na.rm = TRUE)) %>% 
                 mutate(ID = row_number()),
               aes(x = RCP,y = mean_quantity_prct),
               stat = "identity", 
               position = position_dodge2(),
               width = 0.7
               )+ 
      geom_errorbar(data = act_data %>% group_by(SSP, Region,RCP, Product, Year) %>% 
                      summarise(mean_quantity = mean(Quantity, na.rm = TRUE), 
                                SD_quantity = sd(Quantity, na.rm = TRUE),
                                mean_quantity_prct = mean(Quantity_prct, na.rm = TRUE), 
                                SD_quantity_prct = sd(Quantity_prct, na.rm = TRUE)) %>% 
                      mutate(ID = row_number()),
                    aes(x = RCP,ymin=mean_quantity_prct-SD_quantity_prct,
                        ymax=mean_quantity_prct+SD_quantity_prct),
                    width=0.6, size = 0.4, colour="black", alpha=1.2, position = position_dodge(0.7)) +
      scale_fill_manual(values = colour_2070, guide = "none") +
      #facet_grid(.~ RCP)+
      new_scale_colour()+
      geom_point(data = act_data,
                 aes(x=RCP, y = Quantity_prct, colour = DGVM, group = factor(Product)), 
                 shape = 95, size =4, alpha = 1, position = position_dodge(0.7))+
      scale_colour_manual(values = c("#696363", "#E0410B", "#158A11", "#379FA8", "#C419BB"))+
      theme_bw() + 
      #ylim(-3,15)+
      labs(y = "Percentage Change")+
      if(i == 8) {
        theme(
        axis.title = element_blank(),
        axis.text = element_text(size = 11, colour = "black"), 
        axis.text.x = element_text(size = 11, colour = "black"),
        legend.text = element_text(size = 11, colour = "black"), 
        legend.title = element_text(size = 12, colour = "black"), 
        legend.spacing.y = unit(0.3, 'cm'), 
        legend.position="bottom")}  else if (i == 7) {
          
          theme(
            axis.text = element_text(size = 11, colour = "black"), 
            axis.title.y = element_text(colour = "black", size = 12),
            axis.title.x = element_blank(),
            legend.text = element_text(size = 11, colour = "black"), 
            legend.title = element_text(size = 12, colour = "black"), 
            legend.spacing.y = unit(0.3, 'cm'), 
            legend.position="bottom") } else if (i %in% c(1,3,5)) {
              
              theme(
                axis.text.y = element_text(size = 11, colour = "black"), 
                axis.text.x = element_blank(),
                axis.title.x = element_blank(),
                axis.title.y = element_text(colour = "black", size = 12),
                legend.text = element_text(size = 11, colour = "black"), 
                legend.title = element_text(size = 12, colour = "black"), 
                legend.spacing.y = unit(0.3, 'cm'), 
                legend.position="bottom")} else {
          
          theme(
            axis.title = element_blank(),
            axis.text.y = element_text(size = 11, colour = "black"), 
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            legend.text = element_text(size = 11, colour = "black"), 
            legend.title = element_text(size = 12, colour = "black"), 
            legend.spacing.y = unit(0.3, 'cm'),
            legend.position="bottom")
        }
    
    
    
    act_baseline <- baseline_plot %>% filter(Region == unique(data_plot$Region)[i], SSP == "SSP2")
    
    baseline_plot_list[[i]] <- ggplot(data = act_baseline, aes(y = factor(Year, levels = c(2070, 2020)), x = Quantity, fill = Product))+ 
      geom_bar(stat = "identity", position = position_dodge2(), width = 0.7) +
      scale_fill_manual(values =  c(colour_2070))+
      theme_bw() + 
      #annotate("text", x = -Inf, y = -Inf, label = as.character(unique(data_plot$Region)[i]), vjust = -50, hjust = -9) +
      scale_x_continuous(position = "top") +
      labs(y = "Year", x = "Quantity [in ??]")+
      if(i %in% c(1,2)) {
        theme(
          axis.text.y = element_text(size = 11, colour = "black"),
          legend.text = element_text(size = 11, colour = "black"), 
          legend.title = element_text(size = 12, colour = "black"), 
          legend.spacing.y = unit(0.3, 'cm'),
          legend.position="bottom",
          axis.title.x  = element_text(size = 11, colour = "black"),
          axis.title.y = element_blank())

           } else {
            
            theme(
              axis.title = element_blank(),
              axis.text.y = element_text(size = 11, colour = "black"), 
              legend.text = element_text(size = 11, colour = "black"), 
              legend.title = element_text(size = 12, colour = "black"), 
              legend.spacing.y = unit(0.3, 'cm'),
              legend.position="bottom")
            
          }
    
    leg_1<- get_legend(actual_plot_list[[i]])
    leg_2 <- get_legend(baseline_plot_list[[i]])
    
    
    
  quantity_plot_list[[i]] <- ggarrange(baseline_plot_list[[i]],NULL,actual_plot_list[[i]],
                              #labels = c("A",NA, "B"), 
                               heights = c(0.7,0,1),
                               widths = c(1,0,0.1),
                               ncol = 1,
                               align = "v", 
                               legend = "none") 
    
  }
  

leg<- plot_grid(leg_1, NULL,leg_2,
                rel_heights = c(1,0.1,1),
                ncol = 1, 
                align = "v") 


quantity_1 <- ggarrange(quantity_plot_list[[1]],quantity_plot_list[[2]],quantity_plot_list[[3]],quantity_plot_list[[4]],quantity_plot_list[[5]],quantity_plot_list[[6]],quantity_plot_list[[7]],quantity_plot_list[[8]],
          labels = c("A","B","C","D","E","F","G","H"),
          nrow = 4, ncol = 2,
          align = "h")


quantity_2 <- plot_grid(quantity_1,leg,
                rel_heights = c(1,0.1),
                ncol = 1, 
                align = "v") 

quantity_2

```

Price

```{r, warning=FALSE, message=FALSE}
actual_plot_list <- list()
baseline_plot_list <- list()
price_plot_list <- list()

for (i in 1:length(unique(data_plot$Region))) {
  
    act_data <-  data_plot %>% filter(Region == unique(data_plot$Region)[i], SSP == "SSP2") 
    
    actual_plot_list[[i]] <- ggplot(act_data, aes(fill = Product)) + 
      geom_bar(data = act_data %>% group_by(SSP, Region,RCP, Product, Year) %>% 
                 summarise(mean_Price = mean(Price, na.rm = TRUE), 
                           SD_Price = sd(Price, na.rm = TRUE),
                           mean_price_prct = mean(Price_prct, na.rm = TRUE),
                           SD_Price_prct = sd(Price_prct, na.rm = TRUE)) %>% 
                 mutate(ID = row_number()),
               aes(x = RCP,y = mean_price_prct),
               stat = "identity", 
               position = position_dodge2(),
               width = 0.7
               )+ 
      geom_errorbar(data = act_data %>% group_by(SSP, Region,RCP, Product, Year) %>% 
                      summarise(mean_price = mean(Price, na.rm = TRUE), 
                                SD_price = sd(Price, na.rm = TRUE),
                                mean_price_prct = mean(Price_prct, na.rm = TRUE), 
                                SD_price_prct = sd(Price_prct, na.rm = TRUE)) %>% 
                      mutate(ID = row_number()),
                    aes(x = RCP,ymin=mean_price_prct-SD_price_prct,
                        ymax=mean_price_prct+SD_price_prct),
                    width=0.6, size = 0.4, colour="black", alpha=1.2, position = position_dodge(0.7)) +
      scale_fill_manual(values = colour_2070, guide = "none") +
      #facet_grid(.~ RCP)+
      new_scale_colour()+
      geom_point(data = act_data,
                 aes(x=RCP, y = Price_prct, colour = DGVM, group = factor(Product)), 
                 shape = 95, size =4, alpha = 1, position = position_dodge(0.7))+
      scale_colour_manual(values = c("#696363", "#E0410B", "#158A11", "#379FA8", "#C419BB"))+
      theme_bw() + 
      #ylim(-3,15)+
      labs(y = "Percentage Change")+
      if(i == 8) {
        theme(
        axis.title = element_blank(),
        axis.text = element_text(size = 11, colour = "black"), 
        axis.text.x = element_text(size = 11, colour = "black"),
        legend.text = element_text(size = 11, colour = "black"), 
        legend.title = element_text(size = 12, colour = "black"), 
        legend.spacing.y = unit(0.3, 'cm'), 
        legend.position="bottom")}  else if (i == 7) {
          
          theme(
            axis.text = element_text(size = 11, colour = "black"), 
            axis.title.y = element_text(colour = "black", size = 12),
            axis.title.x = element_blank(),
            legend.text = element_text(size = 11, colour = "black"), 
            legend.title = element_text(size = 12, colour = "black"), 
            legend.spacing.y = unit(0.3, 'cm'), 
            legend.position="bottom") } else if (i %in% c(1,3,5)) {
              
              theme(
                axis.text.y = element_text(size = 11, colour = "black"), 
                axis.text.x = element_blank(),
                axis.title.x = element_blank(),
                axis.title.y = element_text(colour = "black", size = 12),
                legend.text = element_text(size = 11, colour = "black"), 
                legend.title = element_text(size = 12, colour = "black"), 
                legend.spacing.y = unit(0.3, 'cm'), 
                legend.position="bottom")} else {
          
          theme(
            axis.title = element_blank(),
            axis.text.y = element_text(size = 11, colour = "black"), 
            axis.text.x = element_blank(),
            axis.ticks.x = element_blank(),
            legend.text = element_text(size = 11, colour = "black"), 
            legend.title = element_text(size = 12, colour = "black"), 
            legend.spacing.y = unit(0.3, 'cm'),
            legend.position="bottom")
        }
    
    
    
    act_baseline <- baseline_plot %>% filter(Region == unique(data_plot$Region)[i], SSP == "SSP2")
    
    baseline_plot_list[[i]] <- ggplot(data = act_baseline, aes(y = factor(Year, levels = c(2070, 2020)), x = Price, fill = Product))+ 
      geom_bar(stat = "identity", position = position_dodge2(), width = 0.7) +
      scale_fill_manual(values =  c(colour_2070))+
      theme_bw() + 
      scale_x_continuous(position = "top") +
      labs(y = "Year", x = "price [in ??]")+
      if(i %in% c(1,2)) {
        theme(
          axis.text.y = element_text(size = 11, colour = "black"),
          legend.text = element_text(size = 11, colour = "black"), 
          legend.title = element_text(size = 12, colour = "black"), 
          legend.spacing.y = unit(0.3, 'cm'),
          legend.position="bottom",
          axis.title.x  = element_text(size = 11, colour = "black"),
          axis.title.y = element_blank())

           } else {
            
            theme(
              axis.title = element_blank(),
              axis.text.y = element_text(size = 11, colour = "black"), 
              legend.text = element_text(size = 11, colour = "black"), 
              legend.title = element_text(size = 12, colour = "black"), 
              legend.spacing.y = unit(0.3, 'cm'),
              legend.position="bottom")
            
          }
    
    leg_1<- get_legend(actual_plot_list[[i]])
    leg_2 <- get_legend(baseline_plot_list[[i]])
    
    
    
  price_plot_list[[i]] <- ggarrange(baseline_plot_list[[i]],NULL,actual_plot_list[[i]],
                              #labels = c("A",NA, "B"), 
                               heights = c(0.7,0,1),
                               widths = c(1,0,0.1),
                               ncol = 1,
                               align = "v", 
                               legend = "none") 
    
  }
  

leg<- plot_grid(leg_1, NULL,leg_2,
                rel_heights = c(1,0.1,1),
                ncol = 1, 
                align = "v") 


price_1<- ggarrange(price_plot_list[[1]],price_plot_list[[2]],price_plot_list[[3]],price_plot_list[[4]],price_plot_list[[5]],price_plot_list[[6]],price_plot_list[[7]],price_plot_list[[8]],
          labels = c("A","B","C","D","E","F","G","H"),
          nrow = 4, ncol = 2,
          align = "h")


price_2 <- plot_grid(price_1,leg,
                rel_heights = c(1,0.1),
                ncol = 1, 
                align = "v") 

price_2

```

**Figure 2 - Standing stock**

Import and rearrange data

```{r, warning = FALSE}
stock <- data.table(rgdx.param("C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/merged.gdx","Forstock1"))

colnames(stock) <- c("SSP","Scenario","Region", "Year","Value")
stock$Scenario <- as.character(stock$Scenario)
stock$Scenario <- ifelse(grepl("^[a-z]", stock$Scenario),
                                      paste0("MC2_", stock$Scenario),
                                      stock$Scenario)

stock <- stock %>%
  filter(Year %in% c("2020", "2050" ,"2070") & Region %in% c("Asia","Europe","Africa", "Oceania", "South America", "North America", "Central America" ))  %>%
  separate(Scenario, into = c("DGVM", "GCM", "RCP"), sep = "_") 

baseline <- stock %>% filter(GCM == "baseline" & Year == 2020)
baseline <- subset(baseline, select = c("SSP", "Region", "Value"))

stock <- left_join(stock, baseline, by = c("SSP", "Region"), suffix = c("", "_baseline"))

```

Calculate Percentage **Question**: No-NPP baseline or just 2020? Not clear from Figure?

```{r, message = FALSE}


stock$percentage <- (stock$Value - stock$Value_baseline) / stock$Value_baseline * 100


```

Create the figure

```{r}
Forstock_plot <- ggplot() + 
  geom_bar(data = stock %>% filter(Year != 2020 & SSP == "SSP2" & RCP %in% c("rcp26", "rcp60", "rcp85")) %>% 
                          group_by(Region, RCP, Year) %>% 
                          summarise(mean_percentage = mean(percentage, na.rm = TRUE), 
                                    SD_percentage = sd(percentage, na.rm = TRUE)), 
           aes(x = Region, y = mean_percentage, fill = Year), 
           stat = "identity", 
           position = position_dodge2()) + 
  geom_errorbar(data = stock %>% filter(Year != 2020 & SSP == "SSP2" & RCP %in% c("rcp26", "rcp60", "rcp85")) %>% 
                          group_by(Region, RCP, Year) %>% 
                          summarise(mean_percentage = mean(percentage, na.rm = TRUE), 
                                    SD_percentage = sd(percentage, na.rm = TRUE)), 
                aes(x = Region, ymin = mean_percentage - SD_percentage, 
                    ymax = mean_percentage+ SD_percentage, fill = Year), 
                position = position_dodge(0.8)) + 
  geom_point(data = stock %>% filter(Year != 2020 & SSP == "SSP2" & RCP %in% c("rcp26", "rcp60", "rcp85")),
                 aes(x=Region, y = percentage, colour = DGVM, group = factor(Year)), 
                 shape = 95, size =4, alpha = 1, position = position_dodge(0.75))+
  facet_grid(.~RCP) + 
  theme_bw()
  
  
  
```

Question: Different values calculated than shown in Figure 2!!

**Figure 3 - Relative Quantity Variation by GDP**

```{r}
GDP <- read.csv("C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/GDPP_Rate.csv")

harvest <- data.table(rgdx.param("C:/Users/maxto/Desktop/Hiwi-Arbeit/FOROM_paper/data/merged.gdx","Forest_Harvest"))

#sort quantity data
harvest <- harvest %>% 
  filter(i5 == "2050") %>% 
  filter(i3 == "T")
colnames(harvest) <- c("SSP","Scenario", "Product","Region", "Year","Value")
harvest$Scenario <- as.character(harvest$Scenario)
harvest$Scenario <- ifelse(grepl("^[[:lower:]a-z]",harvest$Scenario),paste0("MC2_",harvest$Scenario),harvest$Scenario)
  harvest <- harvest %>%
  separate(Scenario, into = c("DGVM", "GCM", "RCP"), sep = "_") 
harvest$Year <- as.character(harvest$Year)

# Baseline creation 
baseline_quantity_variation <- subset(harvest, GCM == "baseline")
baseline_quantity_variation <- baseline_quantity_variation %>% 
  subset(select = -c(Product, RCP))


#subset GDP data to 2050 
GDP <- GDP %>% 
  filter(Year == 2050) %>% 
  select(SSP, Region, Year, GDPP, Growth_rate) %>% 
  mutate(Year = as.character(Year))

GDP$Year <- as.character(GDP$Year)

#combine quantity and GDP data
harvest <- left_join(harvest , GDP %>% select(SSP, Region, Year, GDPP, Growth_rate), by = c("SSP", "Region", "Year"))

#exclude continent 
harvest <- harvest %>% 
  filter(!Region %in% c("Europe","Africa","Asia","Oceania", "North America", "South America", "Central America", "WORLD", "ROW", "United States"))

#assign continent to Region
harvest$continent <- region_to_rcode[as.character(harvest$Region)]

#combine data with baseline
try <- left_join(harvest, baseline_quantity_variation %>% select(-c(DGVM,GCM)), by = c("SSP", "Region", "Year")) %>% 
  rename(value=Value.x, baseline=Value.y)

#calculate relative quantity variation 
try <- try %>% 
  group_by(SSP, Region, GDPP, Growth_rate, continent) %>% 
  summarize(Difference = max(value) - min(value),
            Difference_pct = 100*Difference /mean(baseline),
            .groups = 'drop')
  
```

Create the figure!

```{r, warning=FALSE}
custom_colors <- c("South America" = "#9ee9c4",
                   "Oceania" = "black",
                   "Europe" = "brown",
                   "North America" = "#2a8082",
                   "Asia" = "blue",
                   "Africa" = "#e6ab02")

custom_linetypes <- c("South America" = "solid",
                      "Oceania" = "solid",
                      "Europe" = "dashed",
                      "North America" = "solid",
                      "Asia" = "twodash",
                      "Africa" = "solid")

ggplot(try, aes(x = GDPP, y = Difference_pct)) +
  #geom_point(color = "gray", alpha = 0.3) +
  geom_smooth(aes(color = continent, linetype = continent), method = "lm", se = FALSE) +
  geom_point(data = try, aes(color = continent), alpha = 0.3) +  # Add another geom_point for colored dots
  facet_grid(~SSP) +
  labs(title = "GDP Per Capita vs. Quantity variations by Region",
       x = NULL,
       y = "Quantity Differences %") +
  scale_color_manual(values = custom_colors) +
  scale_linetype_manual(values = custom_linetypes) +
  coord_cartesian(ylim = c(0, 15)) +
  theme_bw() +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) +
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) +
  theme(axis.text.x = element_text(size = 8),
        legend.position = 'bottom',
        legend.direction = "horizontal",
        legend.box = "horizontal",
        legend.key.size = unit(0.5, "cm"),
        legend.key.width = unit(1.2, "cm"),
        legend.text = element_text(size = 8),
        legend.title = element_blank(),
        legend.spacing.x = unit(0.2, "cm"))  # Remove legend for dots
```

## 3. Random Forest Analysis

Doing RFA in Phython

```{python}
import pandas as pd
from sklearn.ensemble import RandomForestRegressor, GradientBoostingRegressor
from sklearn.preprocessing import StandardScaler
from sklearn.inspection import permutation_importance
import matplotlib.pyplot as plt
from matplotlib.backends.backend_pdf import PdfPages

# Read the data from the 'joined1.csv' file --> compiles data from FOROM output
df = pd.read_csv('joined1.csv')

# Get the unique Region values
regions = df['Region'].unique()

# Create a PDF file to save the barplots
pdf_pages = PdfPages('feature_importance_barplots.pdf')

# Create a DataFrame to store all feature importance results
all_importances_df = pd.DataFrame()

# Iterate over each Region and the combined data
for region in list(regions) + ['Combined']:
    if region == 'Combined':
        # Separate features and target variable for combined data
        X = df.drop(['Region', 'Value_Stock'], axis=1)
        y = df['Value_Stock']
    else:
        # Filter the data for the current Region
        region_data = df[df['Region'] == region]
        
        # Separate features and target variable
        X = region_data.drop(['Region', 'Value_Stock'], axis=1)
        y = region_data['Value_Stock']
    
    # Remove rows with NaN values
    X.dropna(inplace=True)
    y = y[X.index]
    
    # Scale the features
    scaler = StandardScaler()
    X_scaled = scaler.fit_transform(X)
    
    # Random Forest
    rf = RandomForestRegressor(n_estimators=100, random_state=42)
    rf.fit(X_scaled, y)
    
    # Sort the feature importances in descending order
    rf_importances = pd.Series(rf.feature_importances_, index=X.columns).sort_values(ascending=False)
    
    # Create a barplot for Random Forest feature importance
    fig, ax = plt.subplots(figsize=(10, 6))
    ax.bar(rf_importances.index, rf_importances.values)
    ax.set_xlabel('Features')
    ax.set_ylabel('Importance')
    ax.set_title(f'Random Forest Feature Importance ({region})')
    ax.tick_params(axis='x', rotation=45)
    plt.tight_layout()
    pdf_pages.savefig(fig)
    
    # Add Random Forest feature importances to the DataFrame
    rf_importances_df = pd.DataFrame({'Feature': rf_importances.index, 'Importance': rf_importances.values})
    rf_importances_df['Region'] = region
    rf_importances_df['Model'] = 'Random Forest'
    all_importances_df = pd.concat([all_importances_df, rf_importances_df], ignore_index=True)
    
    # Gradient Boosting Regressor
    gbr = GradientBoostingRegressor(n_estimators=100, random_state=42)
    gbr.fit(X_scaled, y)
    
    # Sort the feature importances in descending order
    gbr_importances = pd.Series(gbr.feature_importances_, index=X.columns).sort_values(ascending=False)
    
    # Create a barplot for Gradient Boosting Regressor feature importance
    fig, ax = plt.subplots(figsize=(10, 6))
    ax.bar(gbr_importances.index, gbr_importances.values)
    ax.set_xlabel('Features')
    ax.set_ylabel('Importance')
    ax.set_title(f'Gradient Boosting Regressor Feature Importance ({region})')
    ax.tick_params(axis='x', rotation=45)
    plt.tight_layout()
    pdf_pages.savefig(fig)
    
    # Add Gradient Boosting Regressor feature importances to the DataFrame
    gbr_importances_df = pd.DataFrame({'Feature': gbr_importances.index, 'Importance': gbr_importances.values})
    gbr_importances_df['Region'] = region
    gbr_importances_df['Model'] = 'Gradient Boosting'
    all_importances_df = pd.concat([all_importances_df, gbr_importances_df], ignore_index=True)
    
    # Permutation Feature Importance
    perm_importance = permutation_importance(rf, X_scaled, y, n_repeats=10, random_state=42)
    
    # Create a DataFrame with feature names and their permutation importances
    perm_importance_df = pd.DataFrame({'Feature': X.columns, 'Importance': perm_importance.importances_mean})
    perm_importance_df = perm_importance_df.sort_values('Importance', ascending=False)
    
    # Create a barplot for Permutation Feature Importance
    fig, ax = plt.subplots(figsize=(10, 6))
    ax.bar(perm_importance_df['Feature'], perm_importance_df['Importance'])
    ax.set_xlabel('Features')
    ax.set_ylabel('Permutation Importance')
    ax.set_title(f'Permutation Feature Importance using Random Forest ({region})')
    ax.tick_params(axis='x', rotation=45)
    plt.tight_layout()
    pdf_pages.savefig(fig)
    
    # Add Permutation Feature Importances to the DataFrame
    perm_importance_df['Region'] = region
    perm_importance_df['Model'] = 'Permutation'
    all_importances_df = pd.concat([all_importances_df, perm_importance_df], ignore_index=True)

# Close the PDF file
pdf_pages.close()

# Export all feature importances to a single CSV file
all_importances_df.to_csv('all_feature_importances.csv', index=False)
```

creating the figure

```{python}
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns

# Read the CSV data
data = pd.read_csv('all_feature_importances.csv')

# Filter the data for "Gradient Boosting" model
data = data[data['Model'] == 'Gradient Boosting']

# Pivot the data to create a matrix suitable for heatmap
pivot_data = data.pivot_table(index='Region', columns='Feature', values='Importance', aggfunc='mean')

# Create a heatmap
fig, ax = plt.subplots(figsize=(20, 30))  # Increased figure size to accommodate all regions
sns_plot = sns.heatmap(pivot_data, cmap='YlGnBu', annot=True, fmt='.2f', cbar_kws={'label': 'Importance', 'shrink': 0.5}, annot_kws={"fontsize": 16})

ax.set_xlabel('Region', fontsize=16)  # Switch X and Y, and increase font size
ax.set_ylabel('Feature', fontsize=16)  # Switch X and Y, and increase font size
ax.set_title('Factor Importance by Region (Gradient Boosting)', fontsize=28)

plt.xticks(fontsize=20)  # Increase font size for x-tick labels
plt.yticks(rotation=0, fontsize=20)  # Increase font size for y-tick labels and remove rotation

cbar = ax.collections[0].colorbar
cbar.ax.tick_params(labelsize=20)  # Increase font size for colorbar tick labels
cbar.set_label(label='Importance', size=24)  # Increase font size for colorbar label

plt.tight_layout()

# Save the plot as a JPG file
plt.savefig('heatmap.jpg', dpi=300, bbox_inches='tight')

```
